\section{compare\_norms}

\subsection{General}

The compare norms utility calculates distances between sequences
in a dataset using the Smith Waterman distance and the $L_{1}$ and $L_{2}$
distance of k-mer frequency vectors that have been projected on their
principal components.

\subsection{Usage}

The \emph{compare\_norms} tool compares distances between all
sequences in a fasta file in Smith Waterman distance and in a subspace
yielded by PCA. 

\lstset{language=bash,
  caption={Calling the \emph{compare\_norms} tool},
  label=lst-comparenorms-call}
\begin{lstlisting}
compare_norms [fasta] [pca] [dimensions] > [output]
\end{lstlisting}
where the arguments are:
\begin{enumerate}
  \item \emph{fasta} A fasta file containing the sequences to
    calculate the distances in between them.
  \item \emph{pca} A corresponding projection onto principal
    components from k-mers as obtained by the \emph{kmer2pca} tool
    outlined in section \ref{sec-kmer2pca}
  \item \emph{dimensions} The number of projections per sequence to be
    found in the \emph{pca} file.
  \item \emph{output} A file where to output the results.
\end{enumerate}

\subsection{Algorithm}
The Smith Waterman algorithm is used in the same way as pointed out in
section \ref{sec-dbscan-algorithm}. The distance based on the
projections is calculated in $L_1$ and $L_2$ norm.

\subsection{Example}
\lstset{language=bash,
  caption={Example of the \emph{compare\_norms} tool},
  label=lst-comparenorms-example}
\begin{lstlisting}
compare_norms test.fasta test.pca 7 > distances
\end{lstlisting}
The \emph{compare\_norm} tool calculates the distances between the
sequences within the \emph{test.fasta} in Smith Waterman distance and
in $L_1$ and $L_2$ norm in the 7 dimensional PCA subspace.

The distances file contains, per line as first and second integers the
index of the sequences compared. The third position
the Smith Waterman distance, the fourth distance position is the
distance in the PCA subspace in $L_1$ norm and the fifth in $L_2$
norm.
A sample output of looks like this:
\lstset{language=bash,
  caption={Example output of the \emph{compare\_norms} tool},
  label=lst-comparenorms-example-output}
\begin{lstlisting}
i j SW L1 L2
1 0 82.000000 2.546303 6.056748
2 0 122.000000 5.428422 10.829566
2 1 84.000000 4.875662 10.343906
3 0 96.000000 4.237413 8.629056
3 1 88.000000 4.427232 8.065469
3 2 118.000000 5.210451 12.005179
4 0 80.000000 3.201523 7.927563
4 1 72.000000 3.526202 7.919044
4 2 106.000000 6.843701 14.724481
\end{lstlisting}


