\section{find\_satellite}

\subsection{General}

The \emph{find\_satellite} tool is a filtering tool narrowing down a FASTA
dataset of similar sequences. The filter can filter sequences by
length, and by Smith Waterman distance to a template sequence. The
Smith Waterman algorithm is implemented as shown in
\ref{sec-dbscan-algorithm}. This tool only works with nucleotide
sequences. 

\subsection{Usage}
The tool is to be called like:
\lstset{language=bash,
  caption={Calling the \emph{find\_satellite} tool},
  label=lst-findsat-call}
\begin{lstlisting}
find_satellite [fasta-ds] [fasta-target] [size] [size+] [size-] \
  [sw-dist] [n-threads] > [fasta-filtered]
\end{lstlisting}
where the arguments are:
\begin{enumerate}
  \item \emph{fasta-ds} A FASTA dataset of the sequences that should be searched
    for according to the pattern selected by the following arguments.
  \item \emph{fasta-target} A FASTA file holding the target sequence
    from which the Smith Waterman distance will be calculated.
  \item \emph{size} The length of the sequence to be searched for with
    the tolerance intervals given be the subsequent arguments.
  \item \emph{size+} How many nucleotides above the size sequences are
    accepted.
  \item \emph{size-} How many nucleotides below the size sequences are
    accepted.
  \item \emph{sw-dist} The maximum Smith Waterman distance from the
    target sequence to accept sequences.
  \item \emph{n-threads} The number of parallel threads this
    computation may use.
  \item \emph{fasta-filtered} A FASTA file where the sequences that
    met the filtering criteria above are written to. 
\end{enumerate}


\subsection{Example}
\lstset{language=bash,
  caption={Example of the \emph{find\_satellite} tool},
  label=lst-findsat-example}
\begin{lstlisting}
find_satellite test.fasta target.fasta 100 5 3 20 8 > filtered.fasta
\end{lstlisting}
Here the tool searches for sequences in \emph{test.fasta} that are
less then a Smith Waterman distance of 20 away from the sequence
stored in \emph{target.fasta}. Sequences that are between 97 and 105
nucleotides long are accepted. The resulting sequences are stored to
\emph{filtered.fasta}


\subsection{Implementation}
The tools interface is implemented in \emph{find\_satellite.c}. The
mechanism is implemented in \emph{filter.c}.


    
