\section{find\_closest\_sequence\_SW}

\subsection{General}

This tool allows you to find the closest sequence to a known sequence in a
FASTA file, using the Smith Waterman distance as outlined in section
\ref{sec-dbscan-algorithm}.

\subsection{Usage}

The tool can be used like:
\lstset{language=bash,
  caption={Calling the \emph{find\_closest\_sequence\_SW} tool},
  label=lst-closestSW-call}
\begin{lstlisting}
find_closest_sequence_SW [fasta] [fasta-query] [seq-index] [n-threads]
  [fasta-out]
\end{lstlisting}
with the following arguments:
\begin{enumerate}
  \item \emph{fasta} A FASTA file to search for the closest sequence
    in.
  \item \emph{fasta-query} A FASTA file that contains the sequence
    that one wants to find the closest sequence to.
  \item \emph{seq-index} A sequence index that indicates which
    sequence of the \emph{fasta-query} file to use. The first sequence
    has the index 1 in this tool. 
  \item \emph{n-threads} The number of threads that is available to
    the sequence search.
  \item \emph{fasta-out} A FASTA file where the tool will store this
    sequence in. 
\end{enumerate}

\subsection{Example}

\lstset{language=bash,
  caption={Calling the \emph{find\_closest\_sequence\_SW} tool},
  label=lst-closestSW-example}
\begin{lstlisting}
find_closest_sequence_SW test.fasta query.fasta 1 8 result.fasta
\end{lstlisting}
Here the tool finds the closest sequence to the first sequence from
the \emph{query.fasta} file in the \emph{test.fasta} file. The search
might use 8 individual parallel threads. The closest sequence found in
\emph{test.fasta} is written to \emph{result.fasta}.

\subsection{Implementation}
The tool is implemented in \emph{find\_closest\_sequence\_SW.c}.

