\section{reverse\_with\_mask}

\subsection{General}

This tool allows you to reverse selected sequences in a FASTA file. A
mask text file has to be provided to indicate which sequences shall
be reversed. This tool only changes the read direction of the of the
selected sequences. A more elaborate tool is the
\emph{reverse\_complement\_with\_mask} tool found in section \ref{sec-revcomp}.

\subsection{Usage}

The \emph{reverse\_with\_mask} tool can be called like:
\lstset{language=bash,
  caption={Calling the \emph{reverse\_with\_mask\_tool} tool},
  label=lst-revmask-call}
\begin{lstlisting}
reverse_with_mask [fasta] [mask] > [fasta-out]
\end{lstlisting}
with the following arguments:
\begin{enumerate}
  \item \emph{fasta} A FASTA file to reverse sequences in.
  \item \emph{mask} A mask file to indicate the utility which sequences to
    reverse. The file has to contain a line with a number for each
    sequence. If the number matches 1 the sequence corresponding to
    this line in the mask file is reversed in the FASTA output file.
  \item \emph{fasta-out} The FASTA dataset of all sequences
    with those reversed where the mask matches.
\end{enumerate}

\subsection{Example}
Imagine the following FASTA file with three sequences:
\lstset{language={},
  caption={Example of the \emph{reverse\_with\_mask} tool - FASTA file},
  label=lst-revmask-fasta}
\begin{lstlisting}
> sequence_1
ACTG
> sequence_2
ACCT
> sequence_3
GGAT
\end{lstlisting}
As we would like to reverse the second sequence we create the
following mask file:
\lstset{language={},
  caption={Example of the \emph{reverse\_with\_mask} tool - Mask file},
  label=lst-revmask-mask}
\begin{lstlisting}
0
1
0
\end{lstlisting}
Calling the tool as outlined below:
\lstset{language=bash,
  caption={Example of the \emph{reverse\_with\_mask} tool},
  label=lst-revmask-example}
\begin{lstlisting}
reverse_with_mask test.fasta test.mask > out.fasta
\end{lstlisting}
will produce the following FASTA file output:
\lstset{language={},
  caption={Example of the \emph{reverse\_with\_mask} tool - Output FASTA file},
  label=lst-revmask-fastaout}
\begin{lstlisting}
> sequence_1
ACTG
> sequence_2
TCCA
> sequence_3
GGAT
\end{lstlisting}

\subsection{implementation}
The tools interface is implemented in \emph{reverse\_with\_mask.c}.
The code that reverses the sequences resides in \emph{dataset.c}.

