\section{split\_set\_to\_projections} \label{sec-ssproj}

The \emph{split\_set\_to\_projections} tool allows you to get clusterwise
principal component projections from the clustering information stored in
binary \emph{split-set} files obtained by an adaptive clustering run
on a dataset, i.e. obtained from the tool shown in section
\ref{sec-adaptive-clust} and a PCA projection of the same dataset that
can be obtained using the \emph{kmer2pca} tool presented in section
\ref{sec-kmer2pca}.
This tool is especially useful to plot individual clusters together
with a PCA analysis of sequences and hence find sequences
corresponding to clusters in such a diagram. 

\subsection{usage}

The \emph{split\_set\_to\_projections} tool is to be called like:

\lstset{language=bash,
  caption={Calling the \emph{split\_set\_to\_projections} tool},
  label=lst-ssproj-call}
\begin{lstlisting}
split_sets_to_projections [fasta] [pca] [dimensions] [split-set] [projections]
\end{lstlisting}
with the following arguments:
\begin{enumerate}
  \item \emph{fasta} The FASTA file the clustering and hence the
    \emph{split-set} was generated from.
  \item \emph{pca} A projection onto principal components of the
    afformentioned FASTA file, i.e. obtained with the \emph{kmer2pca}
    tool mentioned in section \ref{sec-kmer2pca}.
  \item \emph{dimensions} The number of dimensions of the PCA subspace in the
    afformentioned \emph{pca} file.
  \item \emph{split-set} The binary file that holds the clusters.
  \item \emph{projections} A \emph{path/filename} combination where to put
    the resulting projections of the clusters in PCA space. Each
    cluster will generate a file with an individual number added to
    the name.
\end{enumerate}

\lstset{language=bash,
  caption={Calling the \emph{split\_set\_to\_projections} tool},
  label=lst-ssproj-example}
\begin{lstlisting}
split_sets_to_projections test.fasta test.pca 7 /tmp/s0023 /tmp/s0023proj
\end{lstlisting}
In this case clusters stored in \emph{/tmp/s0023} a binary
\emph{split-set} file are searched in the 7 dimensional subspace
projection onto the principal components of a k-mer representation of
the sequencies found in the \emph{test.fasta} FASTA file. The
projections of each cluster are stored in a seperate file
\emph{/tmp/s0023projX} where X is a number corresponding to each
cluster. 
